---
title             : "PSY 8960-003 Reproducibility Project: Examining the Association Between Neuroticism and Loneliness Among College Students"
shorttitle        : "Reproducibility Project"

author: 
  - name          : "Caroline Armstrong"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "75 East River Parkway, Minneapolis, MN 55455"
    email         : "carolinearmstrong19@gmail.com"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"


affiliation:
  - id            : "1"
    institution   : "Department of Psychology, University of Minnesota"


abstract: |
  [TO COME]
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography: ["references.bib", "r-references.bib"]

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
library(apaTables)
library(dplyr)
library(ggplot2)
library(gtsummary)
library(psych)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(121)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r read-in-data}
data_simulated <- read.csv("../data/neuroticism-loneliness_data-simulated_2025-12-01.csv")
```

[Intro - TO COME]

# Methods

## Data analysis

Descriptive statistics will be computed for demographic variables, including mean and standard deviation for age and, for gender and racial and ethnic identities endorsed, counts and proportions.

I will compute Pearson's *r* for the association between the composite neuroticism and loneliness scores and test the hypothesis that this correlation is significantly ($\alpha$ = .001) greater than *r* = .00 in this sample.

My choice to calculate Pearson's *r* (e.g., versus Spearman's *r*) is based on the fact that the distributions of simulated data for `neuroticism` and `loneliness` are both approximately normal, an assumption of Pearson's *r* (see below for relevant computations and visualizations).

```{r compute-composite-scores}
# Compute composite neuroticism score and composite loneliness score; incorporate these into dataset
data_simulated = data_simulated |> 
  mutate(neur_sum_sim = neur_1_sim + neur_2_sim + neur_3_sim + neur_4_sim + neur_5_sim + neur_6_sim + neur_7_sim + neur_8_sim,
         lic_avg_sim = (lic_1_sim + lic_2_sim + lic_3_sim + lic_4_sim + lic_5_sim + lic_6_sim + lic_7_sim + lic_8_sim + lic_9_sim + lic_10_sim)/10)
```

```{r histogram}
# Visualize distribution of composite neuroticism score and composite loneliness score
hist(data_simulated$neur_sum_sim)
hist(data_simulated$lic_avg_sim)
```

# Results

Using my simulated data, the mean age of participants was `r mean(data_simulated$age)` (SD = `r sd(data_simulated$age)`).

```{r descriptives-age, incude = FALSE}
# Compute mean and SD for age
data_simulated |>
  summarize(mean(age_sim), SD(age_sim)) 
```

Overall, the sample was made up of mostly White women. Descriptive statistics for gender and racial and ethnic identities endorsed are reported in Table 1.

```{r demographics-table}
data_simulated |> tbl_summary(include = c(gender_sim, race_ethn_sim),
                              label = list(gender_sim = "Gender", race_ethn_sim = "Racial and ethnic identities endorsed")) |>
                  modify_caption("**Table 1. Demographic Characteristics of Simulated Data**") 
```

The correlation test revealed that there was not a statistically significant correlation between composite neuroticism and loneliness scores (*p* = 0.9702). See Figure 1.

```{r correlation-test}
# Compute Pearson's r
cor.test(data_simulated$neur_sum_sim, data_simulated$lic_avg_sim,
         alternative = "greater",
         method = "pearson",
         conf.level = 0.95)
```

```{r scatterplot, message=FALSE, warning=FALSE}
# Create scatterplot
ggplot(data = data_simulated, aes(x = neur_sum_sim, y = lic_avg_sim)) +
  geom_point() + 
  theme_apa() +
  xlab("Neuroticism") +
  ylab("Loneliness") + 
  labs(caption = "Figure 1. Loneliness vs. Neuroticism (Simulated Data)") +
  theme(plot.caption = element_text(hjust = 0.1, size = 14))
```

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
